---
output:
  pdf_document: default
  html_document: default
---
```{r}
set.seed(12)
```
### Q1 a)
```{r}
d <- 20
r <- 20000
s <- 5
n <- r * s

k <- d^s

nums <- runif(n)

matrix <- matrix(nums, ncol=s, byrow = TRUE)

N <- numeric(k)

for (i in 1:r) {
  box <- sum((floor(matrix[i, ] * d)) * d^(0:(s-1))) + 1
  N[box] <- N[box] + 1
}

expected_count <- r / k

x_sqr <- sum((N - expected_count)^2 / expected_count)

p = pchisq(x_sqr, df =k-1, lower.tail = FALSE)
print(p)
```
p = 0.07 suggests that the value is uniform enough for our purposes and we
should not reject the null hypothesis

### Q1b)
```{r}
d <- 5
r <- 10000
s <- 5
n <- r * s

k <- d^s
x <- 1
randu = function() {
  x <<- ((2^16 + 3) * x) %% (2^31)
  x/(2^31)
}
nums <- numeric(n)

for (i in 1:n) {
  nums[i] <- randu()
}

matrix <- matrix(nums, ncol=s, byrow = TRUE)

N <- numeric(k)

for (i in 1:r) {
  box <- sum((floor(matrix[i, ] * d)) * d^(0:(s-1))) + 1
  N[box] <- N[box] + 1
}

expected_count <- r / k

x_sqr <- sum((N - expected_count)^2 / expected_count)

p = pchisq(x_sqr, df =k-1, lower.tail = FALSE)
print(p)
```
Since our p value is 0.0049, it is way too small, and thus, the sample is
likely too uniform. Thus, we reject the null hypothesis using d=20, s=5 and
r=20000.

\newpage

### Q2
```{r}
Z <- 7811

nums <- numeric(6)
for (i in 1:6) {
  Z <<- Z^2
  Z <<- sprintf("%08d", Z)
  Z <- as.integer(substr(Z, 3, 6))
  nums[i] <- Z / 10000
}
print(nums)
```

\newpage

### Q3a)
We know that $f(k) \geq 0$, so, we have $c6^{-k} \geq 0$. We also know $\sum_{k=1}^{\infty} c6^{-k} = 1$. With this, we have:

\(
\frac{c}{1-\frac{1}{6}} = 1 \\
c = \frac{5}{6}
\)

### Q3b)
Using the inversion method, we first need to find F:

\(
F(x) = P(W <= x) = \sum_{k \leq x}P(W = k)
F(x) = \sum_{k\leq x}\frac{5}{6}6^{-k} \\
F(x) = \frac{5}{6} \cdot \frac{1 - \big(\frac{1}{6}\big)^{x + 1}}{1 - \frac{1}{6}} = 1 - \Big(\frac{1}{6}\Big)^{x + 1} \\
F^\leftarrow(y) = inf\{x\in\mathbb{R}:\; F(x)\geq y\},\;y\in[0,1] \\
\Big(\frac{1}{6}\Big)^{x + 1} \leq y + 1 \\
(x+1)log\Big(\frac{1}{6}\Big)\leq log(y + 1) \\
x\leq \frac{log(y+1)}{log\big(\frac{1}{6}\big)} - 1 \\
F^\leftarrow(y) = \frac{log(y+1)}{log\big(\frac{1}{6}\big)} - 1
\)

```{R}
nums <- runif(n)

quantile_F <- function(y) {
    return (log(y + 1) / log(1/6) - 1)
}

ans <- sapply(nums, quantile_F)
```

\newpage

### Q4a)

\(
F_X(z) = \mathbb{P}(Z\leq z\;|\;Z\in[a,b]) = \frac{\mathbb{P}(Z\leq z,\;Z\in[a,b])}{\mathbb{P}(Z\in[a,b])} \\
= \frac{\mathbb{P}(Z\in[a,z])}{\mathbb{P}(X\in[a,b])} \\
= \frac{F(z) - F(a)}{F(b) - F(a)}
\)

This is the probability of one sample from Z being in X. The probability of n samples is
\(
n\cdot\frac{F(z) - F(a)}{F(b) - F(a)}
\)

### Q4b)
We can express the CDF of $Z|Z\in[a,b]$ in terms of the CDF F of Z and since the
CDF of a standard normal distribution is continuous and strictly increasing:

\(
F_{Z|Z\in[a,b]} = \frac{F(z) - F(a)}{F(b) - F(a)} \\
F_{Z|Z\in[a,b]}^{-1}(x) = F^{-1}(F(a) + (F(b) - F(a))x)
\)

By the inversion method we can sample n realizations from $Z\;|\;Z\in[a,b]$ as follows:

a) Sample $U_1,\dots,U_n\overset{ind.}{\sim} U(0,1)$.

b) Return $X_j = F^{-1}(F(a) + (F(b) - F(a))x)$ for $j = 1,\dots,n$

### Q4c)
```{r}
n <- 10^7
algo1 <- function(a, b) {
  count <- 0
  while(count < n) {
    num <- rnorm(1)
    if (num >= a && num <= b) {
      count <- count + 1
    }
  }
}

#i)
start_time <- Sys.time()
algo1(-0.3, 0.3)
end_time <- Sys.time()
print(end_time - start_time)

#ii)
start_time <- Sys.time()
algo1(-0.9, 0.5)
end_time <- Sys.time()
print(end_time - start_time)

inverse_f <- function(x, a, b) {
  return (qnorm(pnorm(a) + (pnorm(b) - pnorm(a))*x))
}
algo2 <- function(a, b) {
  u <- runif(n)
  return (sapply(u, inverse_f,a=a, b=b))
}

start_time <- Sys.time()
nums <- algo2(-0.3, 0.3)
end_time <- Sys.time()
print(end_time - start_time)

start_time <- Sys.time()
nums <- algo2(-0.9, 0.5)
end_time <- Sys.time()
print(end_time - start_time)
```

Comparing the CPU times, we can see that for a small range of a and b, the
inverse method is much faster, but for a large range, random sampling is also
a viable method.